Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations

### 2022-01-10 analysis/paper/paper.Rmd
one editorial comment

The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(impact_df$N[impact_df$season == "Fall"]), big.mark = ",")` and during the spring 2021 to be `r prettyNum(unique(impact_df$N[impact_df$season == "Spring"]), big.mark = ",")`. Those estimates translate to linear densities of `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms), big.mark = ",")` which is `r round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater, respectively, than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996]. Although estimated densities for Deadwater Slough are high given the relative size (volume) of the Salmon River at the slough relative to the Columbia River, high CPUE combined with low recapture rates (one in every `r round(nrow(cmr_df) / sum(cmr_df$capture_type == "Recapture", na.rm = T))` fish were recaptures) suggest this area may support exceptionally high predator densities. Further, @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective. Conversely, at Deadwater Slough the maximum depth was approximately 10-15 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach. Our estimates of Northern Pikeminnow areal density are also higher than those reported elsewhere in the Columbia River basin. We estimate the density of Northern Pikeminnow in Deadwater Slough, as fish per hectare, to be `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_hectares), big.mark = ",")` for fall and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_hectares), big.mark = ",")` for spring. For comparison, the only other available estimates of areal density, that the authors are aware of, are 4.4 fish per hectare for John Day Reservoir [@Beamesderfer1991] and 15 fish per hectare in Lake Washington [@Bartoo1977]. Given the dearth of current data available in the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the Columbia River Basin. @Widener2021 has identified a need for more current information on piscine predator populations in the Columbia River.

>>>>

The mean population size of Northern Pikeminnow in the Deadwater Slough during the fall season was estimated to be `r prettyNum(unique(round(impact_df$N[impact_df$season == "Fall"])), big.mark = ",")` and during the spring 2021 to be `r prettyNum(unique(round(impact_df$N[impact_df$season == "Spring"])), big.mark = ",")`. Those estimates translate to linear densities of `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms), big.mark = ",")` 
<!-- need units  -->
which is `r round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / john_day_dens)` and `r round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / john_day_dens)` times greater, respectively, than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r as.character(english(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_rkms / low_columbia_fish_km)))` and `r as.character(english(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_rkms / low_columbia_fish_km)))` times greater than projections for the lower Columbia River [@Beamesderfer1996]. Although estimated densities for Deadwater Slough are high given the relative size (volume) of the Salmon River at the slough relative to the Columbia River, high CPUE combined with low recapture rates (one in every `r round(nrow(cmr_df) / sum(cmr_df$capture_type == "Recapture", na.rm = T))` captured fish were recaptures) suggest this area may support exceptionally high predator densities. Further, @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective. Conversely, at Deadwater Slough the maximum depth was approximately 10-15 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach. Our estimates of Northern Pikeminnow areal density are also higher than those reported elsewhere in the Columbia River basin. We estimate the density of Northern Pikeminnow in Deadwater Slough, as fish per hectare, to be `r prettyNum(round(unique(impact_df$N[impact_df$season == "Fall"]) / ddw_hectares), big.mark = ",")` for fall and `r prettyNum(round(unique(impact_df$N[impact_df$season == "Spring"]) / ddw_hectares), big.mark = ",")` for spring. For comparison, the only other available estimates of areal density, that the authors are aware of, are 4.4 fish per hectare for John Day Reservoir [@Beamesderfer1991] and 15 fish per hectare in Lake Washington [@Bartoo1977]. Given the dearth of current data available in the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the Columbia River Basin. @Widener2021 has identified a need for more current information on piscine predator populations in the Columbia River.
